---
title: "WUP_04"
author: "Danitza Nunez"
date:   "21 SEPT 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
data = read.csv('GSScsv.csv')
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->



We try to predict WORDSUM based on degree. WORDSUM is numerical and DEGREE is categorical, therefore the method used will be numerical categorical. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
par(mar = c(5, 15, 5, 2))
boxplot(WORDSUM~DEGREE,data=data)
unique(data$DEGREE)
```

The X axis represents the level of education and the Y axis represents WORDSUM. Based on the graphical descriptive results, the level of DEGREE with the highest WORDSUM score are the graduates, and the lowest WORDSUM scores are high school and less than high school. Therefore there is a difference in WORDSUM based on education and DEGREE. 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=data)
```


The numerical descriptive results shows that based on DEGREE there is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE. Graduate degrees score a median WORDSUM of 11.While the high school and less than high school scored a median WORDSUM of 7 and even less.  


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfy the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=data)
summary.aov(model1)
```

Based on the inferential results, the P value is 0.000236  smaller than 0.05 therefore we reject the null hypothesis. There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 17, 2, 2))
plot(TukeyHSD(model1),las=2)
```


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
